---
title: "rand"
author: "CZ"
date: "3/25/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(readr)
library(haven)
library(data.table)
library(ggplot2)
library(mosaic)
library(skimr)
#functions
'%ni%' <- Negate('%in%')
```


```{r}
rand <- read_csv("~/Desktop/data/randhrs.csv")
```


Wave 12, which is from 2014. 

# Step 1: Rand data that contains study participants who answered ADL, CESD

We excluded `1-nrow(rand_2014_R_valid)/nrow(rand_2014)` 6.13% data for missing variables for respondents.

We excluded `1-nrow(rand_2014_R_valid)/nrow(rand_2014)` 51.01% data for missing variables for spouses.

We excluded `1-nrow(rand_2014_valid)/nrow(rand_2014))` 53.68% data for missing variables for both respondents and spouses, which left us 2025 observations.
```{r}
#Cohort
table(rand$HACOHORT) 
#Wave 14
table(rand$R12PENINC)
table(rand$R12PENINC,rand$HACOHORT)

#rand_2014   <- read_csv("~/Desktop/data/rand/rand_2014.csv")

rand_2014 <- rand %>%
  select(HHIDPN, HHID, PN, 
         HACOHORT,  #cohort
          R12CESD, S12CESD, #CESD Score, binary level yes or no
         S12AGEY_E,  #Age
          R12ADLA, S12ADLA, #ADL Summary: sum ADLs where Respondent reports any difficulty
          R12OOPMD, S12OOPMD,  #Total expenditure
         RAGENDER,  #gender
         S12SHLT, R12SHLT, #self reported health
         R12HSPTIM,  S12HSPTIM,  # Hospital stays, prv 2 yrs
         R12HSPNIT, S12HSPNIT, # Nights in hosp, prv 2 yrs
         R12NRSNIT, S12NRSNIT,  #Nights in nurs home, prv 2 yrs
         R12WORK, S12WORK, #working for pay
          R12IEARN, S12IEARN#individual earning
        )

#skim(id)
#rename
setnames(rand_2014 ,
         old=c("R12ADLA", "S12ADLA", "R12CESD", "S12CESD", "R12OOPMD", "S12OOPMD", "RAGENDER", "HACOHORT", "S12AGEY_E",
               "R12SHLT", "S12SHLT", 
               "R12HSPTIM", "S12HSPTIM",
               "R12HSPNIT", "S12HSPNIT",
               "R12WORK", "S12WORK",
               "R12IEARN", "S12IEARN"),
         new = c("R_ADL", "S_ADL", "R_CESD", "S_CESD", "R_expense", "S_expense", "sex", "cohort", "age", 
                "R_self_health", "S_self_health",
                "R_hostital_stay","S_hostital_stay",
                "R_nights_hospital","S_nights_hospital",
                "R_work_pay",  "S_work_pay",
                "R_income_i", "S_income_i"))



rand_2014_R_valid <- rand_2014%>%
  filter(!is.na(R_ADL), !is.na(R_CESD), !is.na(R_expense), !is.na(R_self_health), !is.na(R_hostital_stay), !is.na(R_nights_hospital), !is.na(R_work_pay), !is.na(R_income_i))


nrow(rand_2014_R_valid)

# rand_2014_S_valid <- rand_2014%>%
#   filter(!is.na(S_ADL), !is.na(S_CESD), !is.na(S_expense))
# 1-nrow(rand_2014_S_valid)/nrow(rand_2014)

#rand_2014_valid <- rand_2014%>%
#  filter(!is.na(R_ADL), !is.na(R_CESD), !is.na(R_expense),!is.na(S_ADL), !is.na(S_CESD), !is.na(S_expense))
#1-nrow(rand_2014_valid)/nrow(rand_2014)

```


# Helper READ IN
About `1-nrow(helper) /nrow(H14G_HP)` = 50.5% helpers were spouse.  We excluded 49.5% of data.
After filtering out helpers who had invalid helper hour, left us 1243 unique helper in data. 
```{r}
#year: 2014
H14G_HP <- read_sav("/Users/crystalzang/Desktop/data/helper/H14G_HP.sav", encoding = NULL, user_na = FALSE)

###vARIBALE
table(H14G_HP$OPN_SP)
class(H14G_HP$HHID)
H14G_HP$HHID<- as.numeric(H14G_HP$HHID)
H14G_HP$PN<- as.numeric(H14G_HP$PN)
H14G_HP$OPN_SP<- as.numeric(H14G_HP$OPN_SP)

table(H14G_HP$PN)  #helper PN
table(rand_2014_R_valid$PN)  

H14G_HP <- H14G_HP%>%
  filter(!is.na(OPN_SP))%>%
  filter(OPN_SP %ni% c("","811", "812", "821", "831", "832", "841", "822"))%>%
#  mutate(respondent_id= paste(HHID, PN, sep=""), spouse_id=paste(HHID, OPN_SP, sep=""))
  mutate(respondent_id= 1000*HHID + PN, spouse_id=1000*HHID + OPN_SP)

#summary(H14G_HP$respondent_id)

helper <- H14G_HP%>%
  filter(OG069 == "2")

#convert NA value in month/wk/day to 0 
helper<- helper %>%
  mutate(OG070 = coalesce(OG070,0),
         OG071 = coalesce(OG071,0),
         OG072 = coalesce(OG072,0),
         OG073 = coalesce(OG073,0))

#filter out useless values in month/wk/day/hr
helper <- helper %>%
  filter(OG070 %in% 0:31,
         OG071 %in% 0:7,
         OG072 %in% 0:1,
         OG073 %in% 1:24)

#calculate helping hours (assume that one month has 4 weeks and 30 days)
helper <-helper%>%
  mutate(helper_time= OG070*OG073+OG071*4*OG073+OG072*30*OG073)

#helper_unique <- helper[!duplicated(helper$spouse_id), ]
#nrow(helper_unique)

helper_id <- helper%>%
  select(respondent_id, spouse_id, helper_time) %>% 
  mutate(helper="Y")
```


```{r eval=FALSE, include=FALSE}
#plot helper hours
ggplot(data=helper_2014, aes(x=helper_time))+geom_histogram()
count(helper_2014$helper_time==1)

table(helper$OG073)

test <- helper%>%
  group_by(OG070, OG073)%>%
  summarize(N=n())
```


# Join Rand with helper
```{r}
#Join helper with spouse in RAND, indicating spouses who are helpers
class(rand_2014_R_valid$HHIDPN)
class(helper_id$respondent_id)
#rand_2014_R_valid$HHIDPN<-as.character(rand_2014_R_valid$HHIDPN)
helper_id$respondent_id<- as.character(helper_id$respondent_id)

helper <- left_join(rand_2014_R_valid, helper_id, by= c("HHIDPN"= "respondent_id"))%>%
  filter(helper=="Y")%>%
  mutate(helper_status = "helper")

id_group<-helper%>%
  group_by(HHID)%>%
  summarize(N=n())
```

# Type 1 Both helper
```{r}
#check couple
dyad_helper_both<-helper%>%
  group_by(HHID)%>%
  mutate(N=n())%>%
  filter(N==2)
```



# Type 2 Women helper
```{r}
table(helper$sex)

class(helper$spouse_id)
helper$spouse_id<-as.character(helper$spouse_id)

helper_f <- helper %>%
  filter(sex=="2.Female")
nrow(helper_f)


helper_f_id<-helper_f%>%
  select(spouse_id)%>%
  as.data.frame()


helper_f_spouse <- left_join(helper_f_id, rand_2014_R_valid, by= c("spouse_id" = "HHIDPN"))%>%
  mutate(helper_status = "not helper", helper_time = 0)

names(helper_f_spouse)[1]<- "HHIDPN"
helper_f<- helper_f%>%
  select(-spouse_id, -helper)

dyad_helper_f <- rbind(helper_f, helper_f_spouse)

#check couple
 id_group<-dyad_helper_f%>%
   group_by(HHID)%>%
  summarize(N=n())

#Valid dyad HHID 
id_valid_dyad_id <- id_group%>%
  filter(N==2)%>%
  select(HHID)%>%
  mutate(dyad = "Y")

#First join: identify people who are in dyad
dyad_helper_f_join1 <- left_join(rand_2014_R_valid, id_valid_dyad_id, by="HHID")%>%
  filter(dyad == "Y")
#Second join: get information about helper status
dyad_helper_f_id <- dyad_helper_f%>%
  select(HHIDPN, helper_status, helper_time)

dyad_helper_f_join2 <- left_join(dyad_helper_f_join1 , dyad_helper_f_id, by="HHIDPN")

table(dyad_helper_f_join2$helper_status,useNA = "a")

#Cleaned dyad data

#nrow(dyad_helper_f_join2)

#id_group<-dyad_helper_f_final%>%
 # group_by(HHID)%>%
 # summarize(N=n())

#table(dyad_helper_f_final$sex)
#nrow(dyad_helper_f_join2)/2

write.csv(dyad_helper_f_join2, file="~/Desktop/HRS_SEM_2019/data/dyad_helper_f.csv", row.names = F)
```


```{r}
helper_m <- helper %>%
  filter(sex=="1.Male")

helper_m_id<-helper_m%>%
  select(spouse_id)%>%
  as.data.frame()

helper_m_spouse <- left_join(helper_m_id, rand_2014_R_valid , by= c("spouse_id" = "HHIDPN"))%>%
  mutate(helper_status = "not helper", helper_time = 0)

names(helper_m_spouse)[1]<- "HHIDPN"
helper_m<- helper_m%>%
  select(-spouse_id, -helper)

dyad_helper_m <- rbind(helper_m, helper_m_spouse)


#check couple
 id_group<-dyad_helper_m%>%
   group_by(HHID)%>%
  summarize(N=n())

#Valid dyad HHID 
id_valid_dyad_id <- id_group%>%
  filter(N==2)%>%
  select(HHID)%>%
  mutate(dyad = "Y")

#First join: identify people who are in dyad
dyad_helper_m_join1 <- left_join(rand_2014_R_valid, id_valid_dyad_id, by="HHID")%>%
  filter(dyad == "Y")
#Second join: get information about helper status
dyad_helper_m_id <- dyad_helper_m%>%
  select(HHIDPN, helper_status, helper_time)

dyad_helper_m_join2 <- left_join(dyad_helper_m_join1 , dyad_helper_m_id, by="HHIDPN")

table(dyad_helper_m_join2$helper_status,useNA = "a")

#Cleaned dyad data
nrow(dyad_helper_m_join2)

nrow(dyad_helper_m_join2)/2

write.csv(dyad_helper_m_join2, file="~/Desktop/HRS_SEM_2019/data/dyad_helper_m.csv", row.names = F)

```

#Unique Participants
```{r}
id_m<-dyad_helper_m_join2%>%
  select(HHIDPN) #Already unique

id_f<-dyad_helper_f_join2%>%
  select(HHIDPN)#Already unique

#list of unique id
id <- rbind(id_m, id_f)
id_unique<-id[!duplicated(id$HHIDPN), ]
id_ls <- id_unique$HHIDPN


#dyad_all <- rbind(dyad_helper_m_join2, dyad_helper_f_join2)
#In f but not in m 
dyad_diff <- anti_join(dyad_helper_m_join2, dyad_helper_f_join2, by="HHIDPN")

dyad_all<-rbind(dyad_helper_f_join2, dyad_diff)
write.csv(dyad_all, file="~/Desktop/HRS_SEM_2019/data/dyad_all.csv", row.names = F)


#Test
test<-dyad_all%>%
  group_by(HHID)%>%
  summarize(N=n())
table(dyad_all$helper_status,useNA = "a")
```

#Study participants characteristics
```{r}
#Age
helper <- dyad_all%>%
  filter(helper_status == "helper")
respondents <- dyad_all%>%
  filter(helper_status == "not helper")

table(dyad_all$age)
favstats(dyad_all$age)
table(dyad_all$sex)

table(helper$age)
favstats(helper$age)
table(respondents$age)
favstats(respondents$age)

table(respondents$R_ADL,useNA = "a")
dyad_helper_f_resADL0 <- dyad_helper_f_final%>%
  filter(R_ADL==0)

dyad_all
```

